---
title: "Effects of N Addition on AMF Diversity in WI Switchgrass Cropping Systems"
subtitle: "Hancock"
author: "Alden Dirks"
date: "7/26/2019"
output: html_document
---

## Setup

```{r setup, include = FALSE}
rm(list = ls()) # clean up environment for new session
knitr::opts_knit$set(root.dir = '..' ) # set markdown root directory to project root
knitr::opts_chunk$set(echo = FALSE, fig.path = '../figures/', message = FALSE, results = 'hide', warning = FALSE)
```

Load the libraries needed for analysis. `ggpubr` is a package to arrange plots. `lme4` is a standard package for building linear mixed-effects models. `piecewiseSEM` is used to calculate coefficients of determination ($R^2$) for liner mixed-effects models. `tidyverse` is a collection of R packages with the same philosophy and syntax for integrated data science.

```{r libraries, echo = TRUE}
library("ggpubr"); packageVersion("ggpubr") # version 0.2.2
library("lme4"); packageVersion("lme4") # version 1.1.21
library("piecewiseSEM"); packageVersion("piecewiseSEM") # version 2.0.2
library("tidyverse"); packageVersion("tidyverse") # version 1.2.1
```

Load the master dataset and subset for Hancock.

```{r load_data, echo = TRUE}
path_master = file.path("data")
master <- read.table(file.path(path_master, "master.csv"), header = TRUE, check.names = FALSE, sep = ",")
master <- subset(master, site == "Hancock")
tax <- as.matrix(read.table(file.path(path_master, "seqs", "cleaned", "ASVs_taxonomy.csv"), header = TRUE, row.names = 1, check.names = FALSE, sep = ","))
```

## Data Visualization

```{r hancock_pie_charts}
asv_amended <- subset(master, nitrogen == "amended"); asv_amended <- asv_amended[,28:439]
asv_control <- subset(master, nitrogen == "control"); asv_control <- asv_control[,28:439]

asv_amended <- asv_amended[,unname(colSums(asv_amended) > 0)]
asv_control <- asv_control[,unname(colSums(asv_control) > 0)]

tax_amended <- tax[row.names(tax) %in% colnames(asv_amended),]
tax_control <- tax[row.names(tax) %in% colnames(asv_control),]

ggplot(as.data.frame(tax_amended), aes(x = factor(1), fill = Genus)) +
  geom_bar(width = 1) +
  xlab("") + ylab("") +
  coord_polar("y")
ggplot(as.data.frame(tax_control), aes(x = factor(1), fill = Genus)) +
  geom_bar(width = 1) +
  xlab("") + ylab("") +
  coord_polar("y")

ggplot(as.data.frame(tax_amended), aes(x = factor(1), fill = Family)) +
  geom_bar(width = 1) +
  xlab("") + ylab("") +
  coord_polar("y")
ggplot(as.data.frame(tax_control), aes(x = factor(1), fill = Family)) +
  geom_bar(width = 1) +
  xlab("") + ylab("") +
  coord_polar("y")

ggplot(as.data.frame(tax_amended), aes(x = factor(1), fill = Order)) +
  geom_bar(width = 1) +
  xlab("") + ylab("") +
  coord_polar("y")
ggplot(as.data.frame(tax_control), aes(x = factor(1), fill = Order)) +
  geom_bar(width = 1) +
  xlab("") + ylab("") +
  coord_polar("y")
```

```{r hancock_boxplots}
p1 <- ggplot(master, aes(x = nitrogen, y = yield, color = nitrogen)) + 
  geom_boxplot() +
  geom_point() +
  geom_line(aes(group = block), color = "black", alpha = 0.1) + 
  labs(y = "Yield (Mg/ha)", x = "Nitrogen Treatment") +
  theme(legend.position = "none", axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank())
p2 <- ggplot(master, aes(x = nitrogen, y = div_rich, color = nitrogen)) + 
  geom_boxplot() +
  geom_point() +
  geom_line(aes(group = block), color = "black", alpha = 0.1) + 
  labs(y = "Richness", x = "Nitrogen Treatment") +
  theme(legend.position = "none", axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank())
p3 <- ggplot(master, aes(x = nitrogen, y = div_shan, color = nitrogen)) +
  geom_boxplot() + 
  geom_point() +
  geom_line(aes(group = block), color = "black", alpha = 0.1) + 
  labs(y = "Shanon Index", x = "Nitrogen Treatment") +
  theme(legend.position = "none")
p4 <- ggplot(master, aes(x = nitrogen, y = div_simp, color = nitrogen)) +
  geom_boxplot() +
  geom_point() +
  geom_line(aes(group = block), color = "black", alpha = 0.1) + 
  labs(y = "Simpson's Index", x = "Nitrogen Treatment") +
  theme(legend.position = "none")
fig <- ggarrange(p1, p2, p3, p4)
annotate_figure(fig, top = text_grob("Boxplots of Switchgrass Yield and AMF ASV Diveristy by Nitrogen Treatment"))
```

```{r hancock_scatterplots}
p1 <- ggplot(master, aes(x = div_rich, y = yield, color = nitrogen)) + 
  geom_point() + 
  stat_smooth(method = "lm", se = FALSE) + 
  ylim(0, 5) +
  labs(y = "Yield (Mg/ha)", x = "Richness") +
  theme(legend.position = "none")
p2 <- ggplot(master, aes(x = div_shan, y = yield, color = nitrogen)) + 
  geom_point() + 
  stat_smooth(method = "lm", se = FALSE) + 
  ylim(0, 5) + 
  labs(y = "Yield (Mg/ha)", x = "Shannon Index") +
  theme(legend.position = "none", axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks=element_blank())
p3 <- ggplot(master, aes(x = div_simp, y = yield, color = nitrogen)) +
  geom_point() + 
  stat_smooth(method = "lm", se = FALSE) + 
  ylim(0, 5) + 
  labs(y = "Yield (Mg/ha)", x = "Simpson's Index") +
  theme(legend.position = "none")
fig <- ggarrange(p1, p2, p3)
annotate_figure(fig, top = text_grob("Scatterplots of Switchgrass Yield by AMF ASV Diveristy"))
```

## Linear Mixed-effects Models with Likelihood Ratio Test

#### Effects of Nitrogen Treatment on Dependent Variables

```{r hancock_lmer_nitrogen}
# Yield
lmer_nitrogen_yield_null <- lmer(yield ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_yield <- lmer(yield ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_yield)
anova(lmer_nitrogen_yield_null, lmer_nitrogen_yield)

# Richness
lmer_nitrogen_div_rich_null <- lmer(div_rich ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_div_rich <- lmer(div_rich ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_div_rich)
anova(lmer_nitrogen_div_rich_null, lmer_nitrogen_div_rich) # there is a significant effect of nitrogen on ASV richness

# Shannon Index
lmer_nitrogen_div_shan_null <- lmer(div_shan ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_div_shan <- lmer(div_shan ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_div_shan)
anova(lmer_nitrogen_div_shan_null, lmer_nitrogen_div_shan) # there is a significant effect of nitrogen on ASV Shannon index

# Simpson's Index
lmer_nitrogen_div_simp_null <- lmer(div_simp ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_div_simp <- lmer(div_simp ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_yield)
anova(lmer_nitrogen_div_simp_null, lmer_nitrogen_div_simp) # there is a marginally significant effect of nitrogen on ASV Simpson's index

# pH
lmer_nitrogen_ph_null <- lmer(ph ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_ph <- lmer(ph ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_ph)
anova(lmer_nitrogen_ph_null, lmer_nitrogen_ph)

# Soil Percent C
lmer_nitrogen_per_C_null <- lmer(per_C ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_per_C <- lmer(per_C ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_per_C)
anova(lmer_nitrogen_per_C_null, lmer_nitrogen_per_C)

# Soil Percent N
lmer_nitrogen_per_N_null <- lmer(per_N ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_per_N <- lmer(per_N ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_per_N)
anova(lmer_nitrogen_per_N_null, lmer_nitrogen_per_N)

# Soil Carbon:Nitrogen Ratio
lmer_nitrogen_cn_ratio_null <- lmer(cn_ratio ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_cn_ratio <- lmer(cn_ratio ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_cn_ratio)
anova(lmer_nitrogen_cn_ratio_null, lmer_nitrogen_cn_ratio) # there is a significant effect of nitrogen on carbon:nitrogen ratio

# Soil Phosphorus
lmer_nitrogen_p_ppm_null <- lmer(p_ppm ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_p_ppm <- lmer(p_ppm ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_p_ppm)
anova(lmer_nitrogen_p_ppm_null, lmer_nitrogen_p_ppm)

# Catioin Exchange Capacity
lmer_nitrogen_est_cec_null <- lmer(est_cec ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_est_cec <- lmer(est_cec ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_est_cec)
anova(lmer_nitrogen_est_cec_null, lmer_nitrogen_est_cec)

# Coefficient of Determination
model_list <- list(lmer_nitrogen_yield, lmer_nitrogen_div_rich, lmer_nitrogen_div_shan, lmer_nitrogen_div_simp, lmer_nitrogen_ph, lmer_nitrogen_per_C, lmer_nitrogen_per_N, lmer_nitrogen_cn_ratio, lmer_nitrogen_p_ppm, lmer_nitrogen_est_cec)
rsquared(model_list)
```

#### Relationship between Yield and AMF Diveristy and Soil Parameters

```{r hancock_lmer_yield_div_soil}
lmer_yield_null <- lmer(yield ~ nitrogen + (1|block), data = master, REML = FALSE)

# Richness
lmer_yield_rich <- lmer(yield ~ nitrogen + div_rich + (1|block), data = master, REML = FALSE); summary(lmer_yield_rich)
anova(lmer_yield_null, lmer_yield_rich) # there is a significant effect of richness on yield

# Shannon Index
lmer_yield_shan <- lmer(yield ~ nitrogen + div_shan + (1|block), data = master, REML = FALSE); summary(lmer_yield_shan)
anova(lmer_yield_null, lmer_yield_shan) # there is a significant effect of Shannon index on yield

# Simpson's Index
lmer_yield_simp <- lmer(yield ~ nitrogen + div_simp + (1|block), data = master, REML = FALSE); summary(lmer_yield_simp)
anova(lmer_yield_null, lmer_yield_simp) # there is a marginally significant effect of Simpson's index on yield

# pH
lmer_yield_ph <- lmer(yield ~ nitrogen + ph + (1|block), data = master, REML = FALSE); summary(lmer_yield_ph)
anova(lmer_yield_null, lmer_yield_ph) # there is a significant relationship between yield and pH

# Soil Percent C
lmer_yield_C <- lmer(yield ~ nitrogen + per_C + (1|block), data = master, REML = FALSE); summary(lmer_yield_C)
anova(lmer_yield_null, lmer_yield_C)

# Soil Percent N
lmer_yield_N <- lmer(yield ~ nitrogen + per_N + (1|block), data = master, REML = FALSE); summary(lmer_yield_N)
anova(lmer_yield_null, lmer_yield_N) # there is a significant relationship between yield and percent N

# Soil Carbon:Nitrogen Ratio
lmer_yield_CN <- lmer(yield ~ nitrogen + cn_ratio + (1|block), data = master, REML = FALSE); summary(lmer_yield_CN)
anova(lmer_yield_null, lmer_yield_CN) # there is a significant relationship between yield and C:N ratio

# Phosphorous
lmer_yield_P <- lmer(yield ~ nitrogen + p_ppm + (1|block), data = master, REML = FALSE); summary(lmer_yield_P)
anova(lmer_yield_null, lmer_yield_P)

# Cation Exchange Capacity
lmer_yield_CEC <- lmer(yield ~ nitrogen + est_cec + (1|block), data = master, REML = FALSE); summary(lmer_yield_CEC)
anova(lmer_yield_null, lmer_yield_CEC) # there is a significant relationship between yield and cation exchnage capacity

# Coefficient of Determination
model_list <- list(lmer_yield_rich, lmer_yield_shan, lmer_yield_simp, lmer_yield_ph, lmer_yield_C, lmer_yield_N, lmer_yield_CN, lmer_yield_P, lmer_yield_CEC)
rsquared(model_list)
```

#### Relationship between AMF Diversity and Soil Parameters

```{r rhinelander_lmer_div_soil}
## Richness

# Null model
lmer_div_rich_null <- lmer(div_rich ~ nitrogen + (1|block), data = master, REML = FALSE)

# pH
lmer_div_rich_ph <- lmer(div_rich ~ nitrogen + ph + (1|block), data = master, REML = FALSE); summary(lmer_div_rich_ph)
anova(lmer_div_rich_null, lmer_div_rich_ph) 

# Soil Percent C
lmer_div_rich_C <- lmer(div_rich ~ nitrogen + per_C + (1|block), data = master, REML = FALSE)
anova(lmer_div_rich_null, lmer_div_rich_C)

# Soil Percent N
lmer_div_rich_N <- lmer(div_rich ~ nitrogen + per_N + (1|block), data = master, REML = FALSE)
anova(lmer_div_rich_null, lmer_div_rich_N)

# Soil Carbon:Nitrogen Ratio
lmer_div_rich_CN <- lmer(div_rich ~ nitrogen + cn_ratio + (1|block), data = master, REML = FALSE)
anova(lmer_div_rich_null, lmer_div_rich_CN) # significant

# Phosphorous
lmer_div_rich_P <- lmer(div_rich ~ nitrogen + p_ppm + (1|block), data = master, REML = FALSE); summary(lmer_div_rich_P)
anova(lmer_div_rich_null, lmer_div_rich_P) # marginally significant

# Cation Exchange Capacity
lmer_div_rich_CEC <- lmer(div_rich ~ nitrogen + est_cec + (1|block), data = master, REML = FALSE)
anova(lmer_div_rich_null, lmer_div_rich_CEC)


## Shannon Index

# Null model
lmer_div_shan_null <- lmer(div_shan ~ nitrogen + (1|block), data = master, REML = FALSE)

# pH
lmer_div_shan_ph <- lmer(div_shan ~ nitrogen + ph + (1|block), data = master, REML = FALSE); summary(lmer_div_shan_ph)
anova(lmer_div_shan_null, lmer_div_shan_ph)

# Soil Percent C
lmer_div_shan_C <- lmer(div_shan ~ nitrogen + per_C + (1|block), data = master, REML = FALSE)
anova(lmer_div_shan_null, lmer_div_shan_C) # significant

# Soil Percent N
lmer_div_shan_N <- lmer(div_shan ~ nitrogen + per_N + (1|block), data = master, REML = FALSE)
anova(lmer_div_shan_null, lmer_div_shan_N) # there is a significant relationship between div_shan and soil percent N

# Soil Carbon:Nitrogen Ratio
lmer_div_shan_CN <- lmer(div_shan ~ nitrogen + cn_ratio + (1|block), data = master, REML = FALSE)
anova(lmer_div_shan_null, lmer_div_shan_CN)

# Phosphorous
lmer_div_shan_P <- lmer(div_shan ~ nitrogen + p_ppm + (1|block), data = master, REML = FALSE); summary(lmer_div_shan_P)
anova(lmer_div_shan_null, lmer_div_shan_P) # significant

# Cation Exchange Capacity
lmer_div_shan_CEC <- lmer(div_shan ~ nitrogen + est_cec + (1|block), data = master, REML = FALSE)
anova(lmer_div_shan_null, lmer_div_shan_CEC) # marginally significant


## Simpson's Index

# Null model
lmer_div_simp_null <- lmer(div_simp ~ nitrogen + (1|block), data = master, REML = FALSE)

# pH
lmer_div_simp_ph <- lmer(div_simp ~ nitrogen + ph + (1|block), data = master, REML = FALSE); summary(lmer_div_simp_ph)
anova(lmer_div_simp_null, lmer_div_simp_ph)

# Soil Percent C
lmer_div_simp_C <- lmer(div_simp ~ nitrogen + per_C + (1|block), data = master, REML = FALSE)
anova(lmer_div_simp_null, lmer_div_simp_C) # there is a significant relationship between div_simp and soil percent C

# Soil Percent N
lmer_div_simp_N <- lmer(div_simp ~ nitrogen + per_N + (1|block), data = master, REML = FALSE)
anova(lmer_div_simp_null, lmer_div_simp_N) # there is a significant relationship between div_simp and soil percent N

# Soil Carbon:Nitrogen Ratio
lmer_div_simp_CN <- lmer(div_simp ~ nitrogen + cn_ratio + (1|block), data = master, REML = FALSE)
anova(lmer_div_simp_null, lmer_div_simp_CN)

# Phosphorous
lmer_div_simp_P <- lmer(div_simp ~ nitrogen + p_ppm + (1|block), data = master, REML = FALSE); summary(lmer_div_simp_P)
anova(lmer_div_simp_null, lmer_div_simp_P) # significant 

# Cation Exchange Capacity
lmer_div_simp_CEC <- lmer(div_simp ~ nitrogen + est_cec + (1|block), data = master, REML = FALSE)
anova(lmer_div_simp_null, lmer_div_simp_CEC)
```














## Scatterplots

```{r all_sites_scatterplot, include = FALSE}
# biomass
ggplot(master, aes(x = div_rich, y = biomass, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")
ggplot(master, aes(x = div_rar, y = biomass, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")
ggplot(master, aes(x = div_shan, y = biomass, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")
ggplot(master, aes(x = div_simp, y = biomass, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")

# yield
ggplot(master, aes(x = div_rich, y = yield, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")
ggplot(master, aes(x = div_rar, y = yield, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")
ggplot(master, aes(x = div_shan, y = yield, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")
ggplot(master, aes(x = div_simp, y = yield, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")

# height
ggplot(master, aes(x = div_rich, y = height, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")
ggplot(master, aes(x = div_rar, y = height, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")
ggplot(master, aes(x = div_shan, y = height, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")
ggplot(master, aes(x = div_simp, y = height, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")

# lodged
ggplot(master, aes(x = div_rich, y = lodged, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")
ggplot(master, aes(x = div_rar, y = lodged, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")
ggplot(master, aes(x = div_shan, y = lodged, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")
ggplot(master, aes(x = div_simp, y = lodged, color = nitrogen)) + geom_point() + stat_smooth(method = "lm")
```

## Linear Mixed Effects Regression Analysis

#### Biomass

Nitrogen did not significantly affect biomass at Hancock according to LMER ($T_{3.0} = -0.880$, $p = 0.4438$). No metrics of AMF diveristy significantly improvd the model ($p > 0.13$).

```{r predict_biomass}
# LMER
lmer_biomass_null <- lmer(biomass ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_biomass_null)
lmer_biomass_rich <- lmer(biomass ~ nitrogen + div_rich + (1|block), data = master, REML = FALSE); summary(lmer_biomass_rich); anova(lmer_biomass_null, lmer_biomass_rich)
lmer_biomass_rar <- lmer(biomass ~ nitrogen + div_rar + (1|block), data = master, REML = FALSE); summary(lmer_biomass_rar); anova(lmer_biomass_null, lmer_biomass_rar)
lmer_biomass_shan <- lmer(biomass ~ nitrogen + div_shan + (1|block), data = master, REML = FALSE); summary(lmer_biomass_shan); anova(lmer_biomass_null, lmer_biomass_shan)
lmer_biomass_simp <- lmer(biomass ~ nitrogen + div_simp + (1|block), data = master, REML = FALSE); summary(lmer_biomass_simp); anova(lmer_biomass_null, lmer_biomass_simp)
```

#### Yield

Nitrogen did not significantly affect yield at Hancock according to LMER ($T_{3.0} = -1.900$, $p = 0.154$). ASV richness or rarefied richness did not significantly improve the model ($p > 0.45$), however Shannon index and Simpson's index were marginally signiicant ($\chi^2_1 = 2.9567$, $p = 0.08552$) and significant ($\chi^2_1 = 8.7791$, $p = 0.003047$) improvements, respectively.

```{r predict_yield_div, eval = FALSE}
# LMER
lmer_yield_null <- lmer(yield ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_yield_null)
lmer_yield_rich <- lmer(yield ~ nitrogen + div_rich + (1|block), data = master, REML = FALSE); summary(lmer_yield_rich); anova(lmer_yield_null, lmer_yield_rich)
plot(lmer_yield_rich)
lmer_yield_rar <- lmer(yield ~ nitrogen + div_rar + (1|block), data = master, REML = FALSE); summary(lmer_yield_rar); anova(lmer_yield_null, lmer_yield_rar)
plot(lmer_yield_rar)
lmer_yield_shan <- lmer(yield ~ nitrogen + div_shan + (1|block), data = master, REML = FALSE); summary(lmer_yield_shan); anova(lmer_yield_null, lmer_yield_shan)
plot(lmer_yield_shan)
lmer_yield_simp <- lmer(yield ~ nitrogen + div_simp + (1|block), data = master, REML = FALSE); summary(lmer_yield_simp); anova(lmer_yield_null, lmer_yield_simp)
plot(lmer_yield_simp)
```

```{r hancock_scatter}
ggplot(master, aes(x = div_shan, y = yield, color = nitrogen)) + geom_point() + stat_smooth(method = "lm") + labs(x = "ASV Shannon Index", y = "Yield (Mg/ha)")
ggplot(master, aes(x = div_simp, y = yield, color = nitrogen)) + geom_point() + stat_smooth(method = "lm") + labs(x = "ASV Simpson's Index", y = "Yield (Mg/ha)")
```

The following soil parameters were significant improvementes to the null model: percent C, percent N, P, cation exchange capacity. 

```{r predict_yield_soil, eval = FALSE}
# LMER
lmer_yield_null <- lmer(yield ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_yield_null)
lmer_yield_ph <- lmer(yield ~ nitrogen + ph + (1|block), data = master, REML = FALSE); summary(lmer_yield_ph); anova(lmer_yield_null, lmer_yield_ph)
lmer_yield_C <- lmer(yield ~ nitrogen + per_C + (1|block), data = master, REML = FALSE); summary(lmer_yield_C); anova(lmer_yield_null, lmer_yield_C)
plot(lmer_yield_N)
lmer_yield_N <- lmer(yield ~ nitrogen + per_N + (1|block), data = master, REML = FALSE); summary(lmer_yield_N); anova(lmer_yield_null, lmer_yield_N)
plot(lmer_yield_N)
lmer_yield_CN <- lmer(yield ~ nitrogen + cn_ratio + (1|block), data = master, REML = FALSE); summary(lmer_yield_CN); anova(lmer_yield_null, lmer_yield_CN)
lmer_yield_P <- lmer(yield ~ nitrogen + p_ppm + (1|block), data = master, REML = FALSE); summary(lmer_yield_P); anova(lmer_yield_null, lmer_yield_P)
plot(lmer_yield_P)
lmer_yield_cec <- lmer(yield ~ nitrogen + est_cec + (1|block), data = master, REML = FALSE); summary(lmer_yield_cec); anova(lmer_yield_null, lmer_yield_cec)
plot(lmer_yield_cec)
```

```{r rhinelander_scatter_yield_soil}
ggplot(master, aes(x = per_C, y = yield, color = nitrogen)) + geom_point() + stat_smooth(method = "lm") + labs(x = "pH", y = "Yield (Mg/ha)")
ggplot(master, aes(x = per_N, y = yield, color = nitrogen)) + geom_point() + stat_smooth(method = "lm") + labs(x = "Soil Percent N", y = "Yield (Mg/ha)")
ggplot(master, aes(x = p_ppm, y = yield, color = nitrogen)) + geom_point() + stat_smooth(method = "lm") + labs(x = "C:N Ratio", y = "Yield (Mg/ha)")
ggplot(master, aes(x = est_cec, y = yield, color = nitrogen)) + geom_point() + stat_smooth(method = "lm") + labs(x = "Cation Exchange Capacity", y = "Yield (Mg/ha)")
```

#### Diversity

LMER indicated that multiple soil parameters were correlated with ASV Simpson's index, namely percent C, percent N, and P. 

```{r predict_simp_soil, eval = FALSE}
# LMER
lmer_div_simp_null <- lmer(div_simp ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_div_simp_null)
lmer_div_simp_ph <- lmer(div_simp ~ nitrogen + ph + (1|block), data = master, REML = FALSE); summary(lmer_div_simp_ph); anova(lmer_div_simp_null, lmer_div_simp_ph)
lmer_div_simp_C <- lmer(div_simp ~ nitrogen + per_C + (1|block), data = master, REML = FALSE); summary(lmer_div_simp_C); anova(lmer_div_simp_null, lmer_div_simp_C)
plot(lmer_div_simp_C)
lmer_div_simp_N <- lmer(div_simp ~ nitrogen + per_N + (1|block), data = master, REML = FALSE); summary(lmer_div_simp_N); anova(lmer_div_simp_null, lmer_div_simp_N)
plot(lmer_div_simp_N)
lmer_div_simp_CN <- lmer(div_simp ~ nitrogen + cn_ratio + (1|block), data = master, REML = FALSE); summary(lmer_div_simp_CN); anova(lmer_div_simp_null, lmer_div_simp_CN)
lmer_div_simp_P <- lmer(div_simp ~ nitrogen + p_ppm + (1|block), data = master, REML = FALSE); summary(lmer_div_simp_P); anova(lmer_div_simp_null, lmer_div_simp_P)
plot(lmer_div_simp_P)
lmer_div_simp_cec <- lmer(div_simp ~ nitrogen + est_cec + (1|block), data = master, REML = FALSE); summary(lmer_div_simp_cec); anova(lmer_div_simp_null, lmer_div_simp_cec)
```

```{r hancock_scatter_simp_soil}
ggplot(master, aes(x = per_C, y = div_simp, color = nitrogen)) + geom_point() + stat_smooth(method = "lm") + labs(x = "Soil Percent C", y = "AMF ASV Simspon's Index")
ggplot(master, aes(x = per_N, y = div_simp, color = nitrogen)) + geom_point() + stat_smooth(method = "lm") + labs(x = "Soil Percent N", y = "AMF ASV Simspon's Index")
ggplot(master, aes(x = p_ppm, y = div_simp, color = nitrogen)) + geom_point() + stat_smooth(method = "lm") + labs(x = "Soil P (ppm)", y = "AMF ASV Simspon's Index")
```

## Results

Nitrogen did not affect biomass or yield at Hancock. Nothing improved models for biomass, but for yield Simpson's index was a strong predictor. 
