---
title: "Effects of N Addition on AMF Diversity in WI Switchgrass Cropping Systems"
subtitle: "Rhinelander"
author: "Alden Dirks"
date: "7/26/2019"
output: html_document
---

## Setup

```{r setup, include = FALSE}
rm(list = ls()) # clean up environment for new session
knitr::opts_knit$set(root.dir = '..' ) # set markdown root directory to project root
knitr::opts_chunk$set(echo = FALSE, fig.path = '../figures/', message = FALSE, results = 'hide', warning = FALSE)
```

Load the libraries needed for analysis. `ggpubr` is a package to arrange plots. `lme4` is a standard package for building linear mixed-effects models. `piecewiseSEM` is used to calculate coefficients of determination ($R^2$) for liner mixed-effects models. `tidyverse` is a collection of R packages with the same philosophy and syntax for integrated data science.

```{r libraries, echo = TRUE}
library("ggpubr"); packageVersion("ggpubr") # version 0.2.2
library("lme4"); packageVersion("lme4") # version 1.1.21
library("piecewiseSEM"); packageVersion("piecewiseSEM") # version 2.0.2
library("tidyverse"); packageVersion("tidyverse") # version 1.2.1
```

Load the master dataset and subset for Rhinelander. 

```{r load_data, echo = TRUE}
path_master = file.path("data")
master <- read.table(file.path(path_master, "master.csv"), header = TRUE, check.names = FALSE, sep = ",")
master <- subset(master, site == "Rhinelander")
tax <- as.matrix(read.table(file.path(path_master, "seqs", "cleaned", "ASVs_taxonomy.csv"), header = TRUE, row.names = 1, check.names = FALSE, sep = ","))
```

## Data Visualization

```{r rhinelander_pie_charts}
asv_amended <- subset(master, nitrogen == "amended"); asv_amended <- asv_amended[,28:439]
asv_control <- subset(master, nitrogen == "control"); asv_control <- asv_control[,28:439]

asv_amended <- asv_amended[,unname(colSums(asv_amended) > 0)]
asv_control <- asv_control[,unname(colSums(asv_control) > 0)]

tax_amended <- tax[row.names(tax) %in% colnames(asv_amended),]
tax_control <- tax[row.names(tax) %in% colnames(asv_control),]

ggplot(as.data.frame(tax_amended), aes(x = factor(1), fill = Genus)) +
  geom_bar(width = 1) +
  xlab("") + ylab("") +
  coord_polar("y")
ggplot(as.data.frame(tax_control), aes(x = factor(1), fill = Genus)) +
  geom_bar(width = 1) +
  xlab("") + ylab("") +
  coord_polar("y")

ggplot(as.data.frame(tax_amended), aes(x = factor(1), fill = Family)) +
  geom_bar(width = 1) +
  xlab("") + ylab("") +
  coord_polar("y")
ggplot(as.data.frame(tax_control), aes(x = factor(1), fill = Family)) +
  geom_bar(width = 1) +
  xlab("") + ylab("") +
  coord_polar("y")

ggplot(as.data.frame(tax_amended), aes(x = factor(1), fill = Order)) +
  geom_bar(width = 1) +
  xlab("") + ylab("") +
  coord_polar("y")
ggplot(as.data.frame(tax_control), aes(x = factor(1), fill = Order)) +
  geom_bar(width = 1) +
  xlab("") + ylab("") +
  coord_polar("y")
```

At Rhinelander, it appears that yield increased with nitrogen amendment but AMF diveristy was not strongly affected.

```{r rhinelander_boxplots}
p1 <- ggplot(master, aes(x = nitrogen, y = yield, color = nitrogen)) + 
  geom_boxplot() +
  geom_point() +
  geom_line(aes(group = block), color = "black", alpha = 0.1) + 
  labs(y = "Yield (Mg/ha)", x = "Nitrogen Treatment") +
  theme(legend.position = "none", axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank())
p2 <- ggplot(master, aes(x = nitrogen, y = div_rich, color = nitrogen)) + 
  geom_boxplot() +
  geom_point() +
  geom_line(aes(group = block), color = "black", alpha = 0.1) + 
  labs(y = "Richness", x = "Nitrogen Treatment") +
  theme(legend.position = "none", axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank())
p3 <- ggplot(master, aes(x = nitrogen, y = div_shan, color = nitrogen)) +
  geom_boxplot() + 
  geom_point() +
  geom_line(aes(group = block), color = "black", alpha = 0.1) + 
  labs(y = "Shanon Index", x = "Nitrogen Treatment") +
  theme(legend.position = "none")
p4 <- ggplot(master, aes(x = nitrogen, y = div_simp, color = nitrogen)) +
  geom_boxplot() +
  geom_point() +
  geom_line(aes(group = block), color = "black", alpha = 0.1) + 
  labs(y = "Simpson's Index", x = "Nitrogen Treatment") +
  theme(legend.position = "none")
fig <- ggarrange(p1, p2, p3, p4)
annotate_figure(fig, top = text_grob("Boxplots of Switchgrass Yield and AMF ASV Diveristy by Nitrogen Treatment"))
```

It looks like there was a negative relationship between yield and AMF diveristy. 

```{r rhinelander_scatterplots}
p1 <- ggplot(master, aes(x = div_rich, y = yield, color = nitrogen)) + 
  geom_point() + 
  stat_smooth(method = "lm", se = FALSE) + 
  ylim(0, 5) +
  labs(y = "Yield (Mg/ha)", x = "Richness") +
  theme(legend.position = "none")
p2 <- ggplot(master, aes(x = div_shan, y = yield, color = nitrogen)) + 
  geom_point() + 
  stat_smooth(method = "lm", se = FALSE) + 
  ylim(0, 5) + 
  labs(y = "Yield (Mg/ha)", x = "Shannon Index") +
  theme(legend.position = "none", axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks=element_blank())
p3 <- ggplot(master, aes(x = div_simp, y = yield, color = nitrogen)) +
  geom_point() + 
  stat_smooth(method = "lm", se = FALSE) + 
  ylim(0, 5) + 
  labs(y = "Yield (Mg/ha)", x = "Simpson's Index") +
  theme(legend.position = "none")
fig <- ggarrange(p1, p2, p3)
annotate_figure(fig, top = text_grob("Scatterplots of Switchgrass Yield by AMF ASV Diveristy"))
```

## Linear Mixed-effects Models with Likelihood Ratio Test

#### Effects of Nitrogen Treatment on Dependent Variables

According to likelihood ratio tests, nitrogen significantly affected switchgrass yield, AMF richness, and Shannon index. Soil parameters were unaffected by nitrogen addition, although there was a marginally significant effect of nitrogen amendment on carbon:nitrogen ratio and phosphorus. 

```{r rhinelander_lmer_nitrogen}
# Yield
lmer_nitrogen_yield_null <- lmer(yield ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_yield <- lmer(yield ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_yield)
anova(lmer_nitrogen_yield_null, lmer_nitrogen_yield) # there is a significant effect of nitrogen on yield

# Richness
lmer_nitrogen_div_rich_null <- lmer(div_rich ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_div_rich <- lmer(div_rich ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_div_rich)
anova(lmer_nitrogen_div_rich_null, lmer_nitrogen_div_rich) # there is a significant effect of nitrogen on ASV richness

# Shannon Index
lmer_nitrogen_div_shan_null <- lmer(div_shan ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_div_shan <- lmer(div_shan ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_div_shan)
anova(lmer_nitrogen_div_shan_null, lmer_nitrogen_div_shan) # there is a significant effect of nitrogen on ASV Shannon index

# Simpson's Index
lmer_nitrogen_div_simp_null <- lmer(div_simp ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_div_simp <- lmer(div_simp ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_yield)
anova(lmer_nitrogen_div_simp_null, lmer_nitrogen_div_simp)

# pH
lmer_nitrogen_ph_null <- lmer(ph ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_ph <- lmer(ph ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_ph)
anova(lmer_nitrogen_ph_null, lmer_nitrogen_ph)

# Soil Percent C
lmer_nitrogen_per_C_null <- lmer(per_C ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_per_C <- lmer(per_C ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_per_C)
anova(lmer_nitrogen_per_C_null, lmer_nitrogen_per_C)

# Soil Percent N
lmer_nitrogen_per_N_null <- lmer(per_N ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_per_N <- lmer(per_N ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_per_N)
anova(lmer_nitrogen_per_N_null, lmer_nitrogen_per_N)

# Soil Carbon:Nitrogen Ratio
lmer_nitrogen_cn_ratio_null <- lmer(cn_ratio ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_cn_ratio <- lmer(cn_ratio ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_cn_ratio)
anova(lmer_nitrogen_cn_ratio_null, lmer_nitrogen_cn_ratio) # there is a marginally significant effect of nitrogen on carbon:nitrogen ratio

# Soil Phosphorus
lmer_nitrogen_p_ppm_null <- lmer(p_ppm ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_p_ppm <- lmer(p_ppm ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_p_ppm)
anova(lmer_nitrogen_p_ppm_null, lmer_nitrogen_p_ppm) # there is a marginally significant effect of nitrogen on phosphorus

# Cation Exchange Capacity
lmer_nitrogen_est_cec_null <- lmer(est_cec ~ 1 + (1|block), data = master, REML = FALSE)
lmer_nitrogen_est_cec <- lmer(est_cec ~ nitrogen + (1|block), data = master, REML = FALSE); summary(lmer_nitrogen_est_cec)
anova(lmer_nitrogen_est_cec_null, lmer_nitrogen_est_cec)

# Coefficient of Determination
model_list <- list(lmer_nitrogen_yield, lmer_nitrogen_div_rich, lmer_nitrogen_div_shan, lmer_nitrogen_div_simp, lmer_nitrogen_ph, lmer_nitrogen_per_C, lmer_nitrogen_per_N, lmer_nitrogen_cn_ratio, lmer_nitrogen_p_ppm, lmer_nitrogen_est_cec)
rsquared(model_list)
```

#### Relationship between Yield and AMF Diveristy and Soil Parameters

```{r rhinelander_lmer_yield_div_soil}
lmer_yield_null <- lmer(yield ~ nitrogen + (1|block), data = master, REML = FALSE)

# Richness
lmer_yield_rich <- lmer(yield ~ nitrogen + div_rich + (1|block), data = master, REML = FALSE); summary(lmer_yield_rich)
anova(lmer_yield_null, lmer_yield_rich) # there is a significant effect of richness on yield

# Shannon Index
lmer_yield_shan <- lmer(yield ~ nitrogen + div_shan + (1|block), data = master, REML = FALSE); summary(lmer_yield_shan)
anova(lmer_yield_null, lmer_yield_shan) # there is a significant effect of Shannon index on yield

# Simpson's Index
lmer_yield_simp <- lmer(yield ~ nitrogen + div_simp + (1|block), data = master, REML = FALSE); summary(lmer_yield_simp)
anova(lmer_yield_null, lmer_yield_simp) # there is a marginally significant effect of Simpson's index on yield

# pH
lmer_yield_ph <- lmer(yield ~ nitrogen + ph + (1|block), data = master, REML = FALSE); summary(lmer_yield_ph)
anova(lmer_yield_null, lmer_yield_ph) # there is a significant relationship between yield and pH

# Soil Percent C
lmer_yield_C <- lmer(yield ~ nitrogen + per_C + (1|block), data = master, REML = FALSE); summary(lmer_yield_C)
anova(lmer_yield_null, lmer_yield_C)

# Soil Percent N
lmer_yield_N <- lmer(yield ~ nitrogen + per_N + (1|block), data = master, REML = FALSE); summary(lmer_yield_N)
anova(lmer_yield_null, lmer_yield_N) # there is a significant relationship between yield and percent N

# Soil Carbon:Nitrogen Ratio
lmer_yield_CN <- lmer(yield ~ nitrogen + cn_ratio + (1|block), data = master, REML = FALSE); summary(lmer_yield_CN)
anova(lmer_yield_null, lmer_yield_CN) # there is a significant relationship between yield and C:N ratio

# Phosphorous
lmer_yield_P <- lmer(yield ~ nitrogen + p_ppm + (1|block), data = master, REML = FALSE); summary(lmer_yield_P)
anova(lmer_yield_null, lmer_yield_P)

# Cation Exchange Capacity
lmer_yield_CEC <- lmer(yield ~ nitrogen + est_cec + (1|block), data = master, REML = FALSE); summary(lmer_yield_CEC)
anova(lmer_yield_null, lmer_yield_CEC) # there is a significant relationship between yield and cation exchnage capacity

# Coefficient of Determination
model_list <- list(lmer_yield_rich, lmer_yield_shan, lmer_yield_simp, lmer_yield_ph, lmer_yield_C, lmer_yield_N, lmer_yield_CN, lmer_yield_P, lmer_yield_CEC)
rsquared(model_list)
```

#### Relationship between AMF Diversity and Soil Parameters

```{r rhinelander_lmer_div_soil}
## Richness

# Null model
lmer_div_rich_null <- lmer(div_rich ~ nitrogen + (1|block), data = master, REML = FALSE)

# pH
lmer_div_rich_ph <- lmer(div_rich ~ nitrogen + ph + (1|block), data = master, REML = FALSE); summary(lmer_div_rich_ph)
anova(lmer_div_rich_null, lmer_div_rich_ph) # there is a significant relationship between div_rich and pH

# Soil Percent C
lmer_div_rich_C <- lmer(div_rich ~ nitrogen + per_C + (1|block), data = master, REML = FALSE)
anova(lmer_div_rich_null, lmer_div_rich_C)

# Soil Percent N
lmer_div_rich_N <- lmer(div_rich ~ nitrogen + per_N + (1|block), data = master, REML = FALSE)
anova(lmer_div_rich_null, lmer_div_rich_N)

# Soil Carbon:Nitrogen Ratio
lmer_div_rich_CN <- lmer(div_rich ~ nitrogen + cn_ratio + (1|block), data = master, REML = FALSE)
anova(lmer_div_rich_null, lmer_div_rich_CN) # there is a marginally significant relationship between div_rich and C:N ratio

# Phosphorous
lmer_div_rich_P <- lmer(div_rich ~ nitrogen + p_ppm + (1|block), data = master, REML = FALSE)
anova(lmer_div_rich_null, lmer_div_rich_P)

# Cation Exchange Capacity
lmer_div_rich_CEC <- lmer(div_rich ~ nitrogen + est_cec + (1|block), data = master, REML = FALSE)
anova(lmer_div_rich_null, lmer_div_rich_CEC) # there is a significant relationship between div_rich and cation exchnage capacity


## Shannon Index

# Null model
lmer_div_shan_null <- lmer(div_shan ~ nitrogen + (1|block), data = master, REML = FALSE)

# pH
lmer_div_shan_ph <- lmer(div_shan ~ nitrogen + ph + (1|block), data = master, REML = FALSE); summary(lmer_div_shan_ph)
anova(lmer_div_shan_null, lmer_div_shan_ph) # there is a significant relationship between div_shan and pH

# Soil Percent C
lmer_div_shan_C <- lmer(div_shan ~ nitrogen + per_C + (1|block), data = master, REML = FALSE)
anova(lmer_div_shan_null, lmer_div_shan_C)

# Soil Percent N
lmer_div_shan_N <- lmer(div_shan ~ nitrogen + per_N + (1|block), data = master, REML = FALSE)
anova(lmer_div_shan_null, lmer_div_shan_N)

# Soil Carbon:Nitrogen Ratio
lmer_div_shan_CN <- lmer(div_shan ~ nitrogen + cn_ratio + (1|block), data = master, REML = FALSE)
anova(lmer_div_shan_null, lmer_div_shan_CN)

# Phosphorous
lmer_div_shan_P <- lmer(div_shan ~ nitrogen + p_ppm + (1|block), data = master, REML = FALSE)
anova(lmer_div_shan_null, lmer_div_shan_P) # there is a significant relationship between div_shan and P

# Cation Exchange Capacity
lmer_div_shan_CEC <- lmer(div_shan ~ nitrogen + est_cec + (1|block), data = master, REML = FALSE)
anova(lmer_div_shan_null, lmer_div_shan_CEC)


## Simpson's Index

# Null model
lmer_div_simp_null <- lmer(div_simp ~ nitrogen + (1|block), data = master, REML = FALSE)

# pH
lmer_div_simp_ph <- lmer(div_simp ~ nitrogen + ph + (1|block), data = master, REML = FALSE); summary(lmer_div_simp_ph)
anova(lmer_div_simp_null, lmer_div_simp_ph) # there is a significant relationship between div_simp and pH

# Soil Percent C
lmer_div_simp_C <- lmer(div_simp ~ nitrogen + per_C + (1|block), data = master, REML = FALSE)
anova(lmer_div_simp_null, lmer_div_simp_C)

# Soil Percent N
lmer_div_simp_N <- lmer(div_simp ~ nitrogen + per_N + (1|block), data = master, REML = FALSE)
anova(lmer_div_simp_null, lmer_div_simp_N) 

# Soil Carbon:Nitrogen Ratio
lmer_div_simp_CN <- lmer(div_simp ~ nitrogen + cn_ratio + (1|block), data = master, REML = FALSE)
anova(lmer_div_simp_null, lmer_div_simp_CN)

# Phosphorous
lmer_div_simp_P <- lmer(div_simp ~ nitrogen + p_ppm + (1|block), data = master, REML = FALSE)
anova(lmer_div_simp_null, lmer_div_simp_P)

# Cation Exchange Capacity
lmer_div_simp_CEC <- lmer(div_simp ~ nitrogen + est_cec + (1|block), data = master, REML = FALSE)
anova(lmer_div_simp_null, lmer_div_simp_CEC)
```
